---
title: "An implementation of the STAMP protocol."
author: "atb abelew@gmail.com"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    keep_md: false
    mode: selfcontained
    number_sections: true
    self_contained: true
    theme: readable
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
  rmdformats::readthedown:
    code_download: true
    code_folding: show
    df_print: paged
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    width: 300
    keep_md: false
    mode: selfcontained
    toc_float: true
  BiocStyle::html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    keep_md: false
    mode: selfcontained
    toc_float: true
vignette: >
  %\VignetteIndexEntry{error_quantification}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

<style type = "text/css">
body, td {
  font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
 font-size: 16px
}
</style>

```{r options, include = FALSE}
library("stampr")
basedir <- system.file("extdata", package = "stampr")
knitr::opts_knit$set(width = 120,
                     root.dir = basedir,
                     progress = TRUE,
                     verbose = TRUE,
                     echo = TRUE)
knitr::opts_chunk$set(error = TRUE,
                      dpi = 96)
old_options <- options(digits = 4,
                       stringsAsFactors = FALSE,
                       knitr.duplicate.label = "allow")
ggplot2::theme_set(ggplot2::theme_bw(base_size = 10))
rundate <- format(Sys.Date(), format = "%Y%m%d")
##tmp <- sm(loadme(filename = paste0(gsub(pattern = "\\.Rmd", replace = "", x = previous_file), "-v", ver, ".rda.xz")))
##rmd_file <- "03_expression_infection_20180822.Rmd"
library(ggplot2)
```

# Fun with STAMP

This document is intended to step through the STAMP protocol as described in:

1.  "Sequence tag–based analysis of microbial population dynamics"
2.  "Deciphering the landscape of host barriers to Listeria
    monocytogenes infection"
3.  "Systemic infection facilitates transmission
    of Pseudomonas aeruginosa in mice"

And a couple of other papers which I do not have open right now.  I am using a
reference implementation kindly provided by Dr. Lee and Asan which came from one
of the groups listed above, either they haven't told me who or I forgot.
Assuming I receive permission, I will include that implementation here.

# Introduction

There are two broad strokes to the process:

1.  Create a calibration curve showing the relationship between observed CFU and
    tags observed in calibration samples.
2.  Introduce a known set of tags to experimental samples, allow them to
    propagate through the experimental organism, collect DNA samples, quantify
    tags observed at different times/locations, cross reference them against
    the calibration curve, and use that to estimate population bottlenecks
    during the infection process.

From the perspective of the experimental scientist (Dr. Lee and Asan), step 1 is
relatively simple and step 2 is quite onerous.  From the perspective of a computer
nerd, these two vastly different steps are similar and most of the work occurs
in step 1.  This is nice because I have not yet received experimental results.
This document will therefore describe the processing required to go from
sequencer data to a functional calibration curve.  Along the way, I will
introduce some alternatives to the original protocol and provide some example
ways to estimate bottlenecks given artificial experimental input.

With the above in mind, the tasks to process new samples are:

1.  Create a sample sheet, defining the reference, calibration, and experimental
    samples.  I included my copy with this package.
2.  Trim the raw reads.  I used cutadapt through my cyoa preprocessing tool and
    included one output script with this package.  I will leave trimming
    multiple samples as an exercise to the reader.
3.  Count the observed tags.  The authors performed this with qiime.  I will
    show how to do that along with a somewhat more parsimonious method.
4.  Filter the tags and quantify the Nb values.
5.  Create a calibration curve using the quantified tags.
6.  Estimate experimental Nb/CFU using the calibration curve.
7.  Make pretty pictures along the way.

# Installation

The R package may be installed via:

```{r installation, eval = FALSE}
devtools::install_github("abelew/stampr")
```

# Sample metadata

The sample sheet I created is an excel worksheet with one row for each sample.
The most important columns are: "sampleid", "replicate", "type", "CFU",
"index_table", and "qiime_otus".  I think they are mostly self-explanatory, here
are the relevant details:

1.  'type' should be either 'calibration', 'reference', or 'experimental'.
2.  'CFU' should be filled in for the calibration and reference samples.  It will
    be ignored and potentially overwritten for anything labeled 'experimental'.
3.  'index_table' should include filenames for the outputs of my index counter
    script (included in this package) 'count_idx.pl'.
4.  'qiime_otus' should include the filenames produced by qiime's pick_otus.py
    script.

The following block shows the sample sheet from our data.  It is worth noting
that I cheated and used my hpgltools' extract_metadata() function to read it, it
removes any punctuation or shenanigans from the metadata sheet (in case someone
with twitchy fingers added an extraneous space in a filename or something weird).

```{r sample_sheet, results='asis'}
basedir <- system.file("extdata", package = "stampr")
metadata_file <- file.path(basedir, "all_samples.xlsx")
meta <- hpgltools::extract_metadata(metadata_file)
knitr::kable(meta)
```

Note that I put all the raw reads in a series of directories in
'preprocessing'.  I also recompressed every sample and symbolicly linked them to
'r1.fastq.xz'.

# Trim the raw reads

I am not going to include the raw reads in this, but here is the relevant
portion of the script produced by cyoa which performed the trimming for one
sample.  This was one invocation performed on our computing cluster, so I will
remove all the cluster-specific stuff and focus on the important pieces.  Also,
if you look at the version of the script I included there are a few
differences:

1.  I shortened all these paths to relative.
2.  My processor on the cluster by default trims some other adapters which I
    removed for clarity from this.
3.  I added explicit xz invocations to this.  On the cluster I have a series of
    cleanup jobs which do this on multiple nodes.
4.  By default, the less(1) command will not work to read xz/gz/bz2 compressed
    files, and requires the user to first set the LESSOPEN variable, which I
    included at the top.

```{bash trimming, eval = FALSE}
## This was taken from '07ca_r1.sh'
LESSOPEN=| /usr/bin/lesspipe %s
mkdir -p preprocessing/a1/outputs/cutadapt && \
 less r1.fastq.xz | cutadapt -  \
  -a GTCATAGCTGTTT \
  -e 0.1 -n 3 \
  -m 8 -M 42 \
  --too-short-output = preprocessing/a1/outputs/cutadapt/r1_tooshort.fastq \
  --too-long-output = preprocessing/a1/outputs/cutadapt/r1_toolong.fastq \
  --untrimmed-output = preprocessing/a1/outputs/cutadapt/r1_untrimmed.fastq \
  2>outputs/cutadapt.err 1>r1-trimmed_ca.fastq
xz -9e r1-trimmed_ca.fastq preprocessing/a1/outputs/cutadapt/r1_tooshort.fastq \
   preprocessing/a1/outputs/cutadapt/r1_toolong.fastq \
   preprocessing/a1/outputs/cutadapt/r1_untrimmed.fastq
```

# Count observed tags

The original authors used qiime to do this.  I wrote a simple perl script for
it and included it in this package.  The perl script has the small advantage of
being documented and rather fast.

## Qiime OTU quantification

This invocation is from my session on the computing cluster.  Again, this is
just one example.  Also, I did not feel like using the qiime fastq to fasta
converter, I am including it below.  In addition, I found notes in the various
supplemental materials for the previous papers which said that the authors used
99% and 90% identity for pick_otus.  The invocation below uses 90%.

```{bash qiime, eval = FALSE}
module add qiime
fq2fa () {
    paste - - - - < ${1} | cut -f 1,2 | sed 's/^@/>/' | tr "\t" "\n" > ${2}
}

cd preprocessing
start=$(pwd)
for sample in *; do
    cd ${sample}
    fq2fa $(less r1-trimmed_ca.fastq.xz) ${sample}.fasta
    pick_otus.py -i ${sample}.fasta -o $(pwd) -s 0.9
    rm ${sample}.fasta
    xz -9e ${sample}_otus.txt
    cd ${start}
done
```

## Perl tag counting

In contrast, I wrote a short perl script which counts tags without using
ucsearch/blast/etc from qiime.  This has advantages of speed, simplicity, and
it only counts identical tags as opposed to the potential mismatches allowed by
the alignment tools.  The output files are also much simpler 2 column tsv files
with a tag and number on each row.

```{bash perl, eval = FALSE}
cd preprocessing
start=$(pwd)
for sample in *; do
    cd ${sample}
    ../count_idx.pl r1-trimmed_ca.fastq.xz
    xz -9e idx_count.txt
    cd ${start}
done
```

## Look for yourself

If you wish to try it for yourself, I included archives of the trimmed reads,
the outputs from pick_otus.py and the outputs from count_idx.pl.

### Extract the trimmed reads

```{r extract_reads}
untarred <- utils::untar(tarfile = system.file("extdata/trimmed_reads.tar", package = "stampr"))
```

### Extract the otus

```{r extract_otus}
untarred <- utils::untar(tarfile = system.file("extdata/otus.tar", package = "stampr"))
```

### Extract the tag counts

```{r extract_counts}
untarred <- utils::untar(tarfile = system.file("extdata/tag_counts.tar", package = "stampr"))
```

At this point, the current working directory should have files in it for the
trimmed reads, the qiime otus, and the counts from my perl script.  For the
following steps, I will begin by performing each task as the original authors
did them.  I will then show how to perform them using this package.

# Tag quantification

## Original with slight changes

This does not work without the qiime output files from the authors, which I
don't have.  I hypothesize that they changed the format of the output files, but
I am not completely certain about how they were changed, so I am just going to
leave it here.

```{r tally, eval = FALSE}
## This R script will take the output files containing representative STAMP barcode sequences and
## their individual abundance in the sequencer output from all samples  that
## you sequenced and copies them in a single .csv file in the format that is
## expected from the Nb estimation script

output_file <- "qiime_tally.csv"
## files <- list.files(txt_dir)
files <- meta[["qiimeotus"]]
names <- files
v.length <- rep(NA, length(names))
for (f in 1:length(files)) {
  file <- files[f]
  ##mtrx <- read.table(file, sep = "\t", header = FALSE, fill = TRUE)
  ##colnames(mtrx) <- c(paste0(files[f], "_seq"), files[f])
  eval(parse(text = paste("Data = as.matrix(read.table('",file,"', fill = TRUE, sep='\t'))",sep = "")))
  colnames(Data) <- c(paste(names[f],"_seq", sep = ""), names[f])
  name <- paste("Matrix", names[f], sep = "")
  assign(name, Data)
  name <- paste("Matrix", names[f], sep = "")
  assign(name, Data)
  v.length[f] <- nrow(Data)
}
max <- max(v.length) + 1
fill.up <- max - v.length

FillUpMatrix <- matrix(data = NA, nrow = fill.up[1], ncol = 2)
eval(parse(text = paste("OutputMatrix = rbind(Matrix", names[1],", FillUpMatrix)", sep = "")))
for(ff in 2:length(files)) {
  eval(parse(text = paste("MatrixX = Matrix", names[ff], sep = "")))
  FillUpMatrix <- matrix(data = NA, nrow = fill.up[ff], ncol = 2)
  Matrix <- rbind(MatrixX, FillUpMatrix)
  OutputMarix <- cbind(OutputMatrix, Matrix)
}
write.csv(OutputMatrix, file = outputfilename)
```

All of the functions in this package require a metadata dataframe as input.
Thus, meta will be the first parameter for pretty much every function.

## My method using actual qiime output files

In my current invocation qiime does not return the tag sequences but instead
calls them denovoXXX because it was run in its denovo mode.  In terms of result,
this is a little weird but not relevant.

```{r read_qiime_otus}
qiime_tags <- read_qiime_otus(meta, otu_column = "qiimeotus", trimmed_column = "trimmedfastq")
head(qiime_tags[["modified_metadata"]])
meta <- qiime_tags[["modified_metadata"]]

## Number of reads / tag
plot_read_density(qiime_tags)
```

## My method using the perl tag counts

In the following block, I essentially repeat the above using tags as
defined by my simplified perl script.

```{r read_perl_tags}
perl_tags <- read_tags(meta, index_column = "indextable")
summary(perl_tags)
## read_idx returns a list with 3 elements:
##  the input metadata with some new columns.
##  the counts in a long format for easier plotting.
##  the counts in a wide format, which is easier to work with.
head(perl_tags[["modified_metadata"]])
## The new elements in the metadata include:
##  reads:  how many reads were used in this counting.
##  observed: how many tags were observed.
##  passed_cutoff: how many tags passed our simple cutoff for 'real'.
##  rejected_cutoff: passed + rejected = observed
##  max: What is the maximum number of reads for 1 tag.  Thus for a1, we see that one tag sucked up
##       almost half of all the reads.

## Given the above, we can trivially plot how many reads/tag there are for each sample:
plot_read_density(perl_tags)
```

## The original qiime tally file

Note that the original method has a slightly different format.

```{r original_qiime_tally}
original_tag_file <- system.file("extdata/original_qiime_tally.csv", package = "stampr")
original_tags <- read.csv(original_tag_file)
head(original_tags)
```

# Filter the tags

There are lots of ways one may choose to filter the tags to get the set of
'real' tags.  Ideally, one would take the tags observed in the reference samples
and remove anything not in it.  Sadly, in our current data, there are far far
fewer tags in the references than the calibration samples, so I have not
implemented that.  Conversely, one might choose to remove tags observed in < x
samples.  I haven't implemented that yet either.  I did, however, cpm() the
matrix and remove anything less than 1x the number of samples (which is configurable).

```{r filter_tags}
new_meta <- perl_tags[["modified_metadata"]]
current_tags <- perl_tags[["sample_counts"]]
filtered_tags <- filter_tags(perl_tags)
summary(filtered_tags)
head(filtered_tags[["modified_metadata"]])
## Thus the new column 'passed_aggressive_cpm'
```

# Quantify the Nb values.

## The original with slight changes

The following block contains the original code with slight modifications and
produces the same Nb values as provided.

```{r quantify_nb_original, eval = FALSE}
inputdata <- "../../original/BigMatAll_STAMP_CAUTI_calcurve1_working.csv"
### read input sequence data file
## Data = as.matrix(read.csv(inputdata,row.names = 1))
Data <- read.csv(inputdata)
rownames(Data) <- make.names(Data[["X"]], unique = TRUE)
Data[["X"]] <- NULL

#BigMatAll = DataII[,2:ncol(DataII)]
BigMatAll <- Data

### USER ### define output filename
name <- "New_EstimateMatrix_CAUTI_Calcurve1"

### shows colnames and col # to make it easier to identify the reference samples
### and the data samples
cbind(1:ncol(BigMatAll), colnames(BigMatAll))

### should a calibration curve be used to normalize the data? TRUE/FALSE
usecalibrationcurve <- FALSE

### USER ### If calibration curve is used, give path to calibration data
### ("CalibrationLookup")
if (usecalibrationcurve) {
  load("/Users/kerbachta/Desktop/STAMP/calcurve3/CalibrationLookup.Rdata")
}

### USER ### define which columns of BigMatAll contain real read numbers and not
### meta-data like binaries, average count number or absolute presence in
### samples. Usually column 1 to 24 (1:24) contain real read numbers.
v.reads <- 1:23

### USER ### define which columns of BigMatAll contain reference samples
### (usually the inoculum)? Define at least two columns e.g. 1:2 or c(1,2) would
### mean column 1 and column 2 of BigMatAll contain reference samples.
## v.reference <- 21:23
v.reference <- c(8, 15, 23)

### USER ### define which columns of BigMatAll contain data samples (that should
### be the columns of BigMatAll that contain real read numbers minus the
### reference columns)
## v.bottlenecksamples <- 1:20
v.bottlenecksamples <- c(1:7, 9:14, 16:22)

### define threshold for inocula (e.g. sequence must be present at least once)
v.threshold <- which(rowSums(BigMatAll[, v.reference], na.rm = TRUE) > 0)

### choose real reads
ThresholdMatrix <- BigMatAll[v.threshold, v.reads]

## Why was this not done first?
### NA= 0 (not present)
ThresholdMatrix[is.na(ThresholdMatrix) == TRUE] <- 0

### Create matrix with references only
ReferenceMatrix <- ThresholdMatrix[, v.reference]
head(ReferenceMatrix)

### Define FrequenzMatrix
fReferenceMatrix <- matrix(nrow = nrow(ReferenceMatrix),
                           ncol = ncol(ReferenceMatrix),
                           dimnames = list(rownames(ReferenceMatrix),
                                         colnames(ReferenceMatrix)))

### Fill FrequenzMatrix with frequencies
for(x in 1:ncol(fReferenceMatrix)) {
    fReferenceMatrix[, x] <- ReferenceMatrix[, x] / sum(ReferenceMatrix[, x])
}

### Average frequencies
fTotalInoculum <- rowMeans(fReferenceMatrix, na.rm = TRUE)

### make matrix that connects bottleneckestimate to columnname
EstimateMatrix <- matrix(nrow = 5, ncol = length(v.bottlenecksamples))
colnames(EstimateMatrix) <- colnames(BigMatAll[, v.bottlenecksamples])
rownames(EstimateMatrix) <- c("Nb","lower CI","Nb'_median","upper CI","# sequences")

### define expected counts and frequencies
expectedCounts <- rowMeans(ReferenceMatrix, na.rm = TRUE)
expvec <- fTotalInoculum[expectedCounts > 0]
inoculumSize <- sum(ReferenceMatrix, na.rm = TRUE)

### loop for estimating Ne for each sample
for (i in 1:length(v.bottlenecksamples)) {
  bnvec <- ThresholdMatrix[, v.bottlenecksamples[i]]
  bnvec <- bnvec[expectedCounts > 0]
  sampleSize <- sum(bnvec)
  binomialNenner <- expvec * (1 - expvec)
  measured <- bnvec / sum(bnvec)
  var <- (measured - expvec) ^ 2
  binomial <- as.numeric(var) / as.numeric(binomialNenner)
  BSize <- 1 / (mean(binomial) - 1 / sampleSize - 1 / inoculumSize)
  EstimateMatrix[1, i] <- BSize
  Log10BSize <- round(log10(BSize), 2)
  EstimateMatrix[5, i] <- sampleSize
### this part of the code compares to the lookupmatrix
  if (usecalibrationcurve) {
    if (Log10BSize < max(as.numeric(rownames(LookUpMatrix)))) {
      if (usecalibrationcurve) {
        EstimateMatrix[2:4, i] <- 10 ^ LookUpMatrix[as.character(Log10BSize), ]
      }
    }
  }
}

head(EstimateMatrix)
tmp <- t(EstimateMatrix)
meta[["Nb_original"]] <- 0
for (l in 1:nrow(tmp)) {
  row <- tmp[l, ]
  name <- rownames(tmp)[l]
  name <- tolower(gsub(x = name, pattern = "\\.txt", replacement = ""))
  meta[name, "Nb_original"] <- row[["Nb"]]
}
```

## My version from the index counts

```{r my_nb_idx}
my_nb <- calculate_nb(perl_tags, metadata_column = "MyNb")

qiime_nb <- calculate_nb(qiime_tags, metadata_column="qiimenb")
```

## Calculate Nb after top-n tag filtering.

```{r nb_filtered}
topn_filtered <- filter_topn_tags(my_nb, topn = 600)
topn600_nb <- calculate_nb(topn_filtered, metadata_column = "Nb600")
top500_filtered <- filter_topn_tags(topn600_nb, topn = 500)
topn500_nb <- calculate_nb(top500_filtered, metadata_column = "Nb500",
                           write = "modified_metadata-20210105.xlsx")
```

# Create Calibration curve

```{r plot_calibration}
cali_curve <- plot_calibration(qiime_nb)
cali_curve$plot
cali_curve$lm

cali_curve_top500 <- plot_calibration(topn500_nb, transform_x = "log10", transform_y="log10")
cali_curve_top500$plot
cali_curve$lm

cali_curve <- plot_calibration(my_nb, transform_x = "log10", transform_y = "log10")
cali_curve$plot
cali_curve$lm
hpgltools::pp(file = "conservative_calibration_curve.png", image = cali_curve$plot)

topn500_nb_nolast <- exclude_samples(topn500_nb, max_dilution=1e-02)
cali_curve_log10 <- plot_calibration(topn500_nb_nolast, transform_x = "log10", transform_y = "log10")
cali_curve_log10$lm
cali_curve_log10$plot
hpgltools::pp(file = "conservative_calibration_curve_log10_nolastdilution.pdf", image = cali_curve_log10$plot)
```

5.  Create a calibration curve using the quantified tags.
6.  Estimate experimental Nb/CFU using the calibration curve.
7.  Make pretty pictures along the way.
